Characterization of gaps in directly bonded Si compound optics

### 2014-03-20 AO_bondedSi.tex
some section rearrangement and edits for clarity

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.95\columnwidth]{figs/VS20fineGapCrop.pdf}}
\caption{Veeco optical profiler surface plot and histogram for part VS20 in the fine pattern section\label{figVS20pattern}.  The four inch diameter VS20 is comprised of two thick silicon pucks.  We patterned three meshes with coarse, medium, and fine boxes.  The depth was 95 nm as indicated by the well-separated peaks in the Veeco profilometry histogram.  The Veeco was operated in PSI mode.  The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.  The technique can reveal the axial extent and fill factor of an embedded gap.}
\end{figure}

@@

\section{Transmission measurement theory}
\label{secTheory}

@@

Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.  The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.  The Fresnel reflection is accordingly about 30\% per surface.  For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.  Figure X shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{fig} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}, and extrapolated past their upper temperature bound to estimate values near 373 K.  The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.  The minuscule temperature dependence is largely negligible, and for the rest of the paper we assume room temperature refractive index.}
\end{figure}

The Si-air\footnote{We imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.} interfaces of the in the bond gap make a Fabry-P\`erot etalon\cite{2007fuph.book.....S}.  The transmission through a Fabry-Perot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.  We assume transmission is at normal incidence, the index of the gap is 1.0, and the material is at room temperature.  We assume Si is nonabsorbing longward of $\lambda$ = 1150 nm.  We compute the reflectivity for an Si-air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}.  The computation is simply Fresnel's law at normal incidence:

$$
R = \frac{(n-1)^2}{(n+1)^2}
$$

The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.  The modest wavelength dependence is shown in the bottom curve of the figure below.  The coefficient of Finesse\cite{2007fuph.book.....S} encapsulates the Fabry-P\`erot etalon's dependence on the reflectivity.  We compute a coefficient of Finesse in the range of about 2.7$-$2.9 depending on the wavelength.

\subsection{Consideration of multiple incoherent reflections}
\label{secTheory}
Multiple incoherent reflections occur within the Si substrate.  Treatment for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.  The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.  I adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.  Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.  The details are worked out in Appendix \ref{sec:Append-IMRTMM}.  We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.  The DSP Si puck with no air gap has a transmission equal to:

@@

which has an average value of about 53\%, as shown in the top curve of the Figure above.  Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.  If the Si wafer path length is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.  We work out this behavior in the Appendix section X and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.

For the case of a pair of bonded Si pucks with a gap thickness $d$, we normalize the transmission with a gap to the transmission with no gap to isolate the effect of the gap.  We call this normalized transmission the etalon transmission $T_{e}$.  It is derived in Equation \ref{eqn:Tetalon} in the Appendix:

@@

\section{Substrate preparation}
It is important to clean the wafers to minimize the particle density, since particles cause gaps.  The substrates were silicon wafers.  The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler .  The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.  We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.  The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.  There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.  We measured the large scale surface flatness of the pucks.  The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.  We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.

The detailed bonding procedure will be chronicled in a future paper.  Briefly, we prepared the substrates with standard cleaning procedures.  We used MHz frequency oxygen plasma ashing.  We soaked the wafers in DI water then dried with an N$_2$ gun.  We pressed the wafers together from the center.

>>>>

\begin{figure}[htbp]
\centerline{\includegraphics[width=1.0\columnwidth]{figs/VS20fineGapCrop.pdf}}
\caption{Veeco optical profiler surface plot and histogram for part VS20 in the fine pattern section\label{figVS20pattern}.  The four inch diameter VS20 is comprised of two thick silicon wafers.  We patterned three meshes with coarse, medium, and fine boxes, each with 50\% fill factors.  The depth was 95 $\pm$ 5 nm as indicated by the well-separated peaks in the Veeco profilometry histogram.  The Veeco was operated in PSI mode.  The purpose for these patterns was to have gaps of known dimensions against which we could test the optical metrology technique we describe in this paper.  The technique can reveal the axial extent and fill factor of an embedded gap.  }
\end{figure}

@@

\section{Substrate preparation}
It is important to clean the surfaces before bonding to minimize the particle density.  The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.  The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.  We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.  The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.  There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.  We measured the large scale surface flatness of the pucks.  The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.  We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.

The detailed bonding procedure will be chronicled in a future paper.  Briefly, we prepared the substrates with standard cleaning procedures of solvents in a megasonic.  We then applied MHz frequency oxygen plasma ashing.  We soaked the wafers in DI water then dried with an N$_2$ gun.  We pressed the wafers together from the center to the outside with normal human force.

\section{Si$-$air$-$Si Fabry P\`{e}rot measurement theory}
\label{secTheory}

@@

\subsection{The Si$-$air$-$Si cavity is a low-Finesse Fabry-P\`{e}rot etalon}

Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si-Air interface.  The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.  The Fresnel reflection is accordingly about 30\% per surface.  For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.  Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}, and extrapolated past their upper temperature bound to estimate values near 373 K.  The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.  The minuscule temperature dependence is largely negligible, and for the rest of the paper we assume room temperature refractive index.}
\end{figure}

The Si$-$air$-$Si cavity is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.  Note that we imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.  The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.  We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.  We assume Si is nonabsorbing longward of $\lambda$ = 1150 nm.  We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.  The computation is simply Fresnel's law at normal incidence:

$$
R = \frac{(n-1)^2}{(n+1)^2} \\
F \equiv \frac{4R}{(1-R)^2}
$$

The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.  We've defined the coefficient of Finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.  The coefficient of Finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.

\subsection{Consideration of multiple incoherent reflections}
\label{secTheory}
Multiple incoherent reflections occur within the Si substrate.  Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.  The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.  We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.  Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.  The details are worked out in Appendix \ref{sec:Append-IMRTMM}.  We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.  The DSP Si puck with no air gap has a transmission equal to:

@@

which has an average value of about 53\%, as shown in the top curve of the Figure above.  Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.  If the Si wafer path length is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.  This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.
%We work out this behavior in the Appendix and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.

Consider a pair of bonded Si pucks with a gap thickness $d$. We derive the absolute transmission bonded Si wafers with a gap in Equation \ref{eqn:Tetalon} in the Appendix.  We normalize the absolute transmission by the transmission with no gap to isolate the effect of the gap.  We call this normalized transmission the etalon transmission $T_{e}$, since it isolates the effect of the etalon on the transmission.

@@
